---
title: "Assignment_I"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



#### Section 1: Probability : Total points 50


Q1. Which of the following numbers cannot be probability? Explain why.     

a) -0.0001
b) 0.05
c) 3.415
d) 20%
e) 1

(Ans) 1.  (a) and (c) cannot be probability of any event because the probability of an event is always non-negative and lies between 0 and 1.

Q2. A card is drawn randomly from a deck of ordinary playing cards. The game rules are that you win if the card is a spade or an ace. What is the probability that you will win the game?

(Ans) 2. 3/52 + 13 /52 = 16/52 or 0.308 or 30.8% chance of wininng the game.

Q3. An urban hospital has a 20% mortality rate on average for admitted patients. If on a particular day, 17 patients got admitted, what are:

a) the chances that exactly 7 will survive?
```{r}
#A binomial distribution that selects exactly 7 patients out of 17 to survive with a rate of 0.8. (or mortality rate = 0.2)
p_7_surive <-  dbinom(x = 7, size = 17, prob = 0.8)
print(p_7_surive)
```

b) the chances that at least 15 patients will survive?    
```{r}
# P(at least 15 survive) = P(Exactly 15 survive) + P(Exactly 16 survive) + P(Exactly 17 survive)
p15 <- dbinom(x = 15, size = 17, prob = 0.8)
p16 <- dbinom(x = 16, size = 17, prob = 0.8)
p17 <- dbinom(x = 17, size = 17, prob = 0.8)
pAtleast15 <- p15 + p16 + p17
print(pAtleast15)
```


Q4. Let F and G be two events such that P(F) is 0.4, P(G) is 0.8. F and G are independent events. Fill in the remaining elements of the table.

Table    | $G$ | $\bar{G}$ | Marginal
---------|-----|-----------|----------
$F$      |0.32 |  0.08     | 0.40 
$\bar{F}$|0.48 |  0.12     | 0.60
Marginal |0.80 |  0.20     |    1


Q5. Let F and G be two events such that P(F) is 0.2, P(G) is 0.7. Now, the conditional probability P(G|F) is given as 0.4. Fill in the remaining elements of the table.

Table    | $G$ | $\bar{G}$ | Marginal
---------|-----|-----------|----------
$F$      |0.08 |    0.12   |   0.2
$\bar{F}$|0.62 |    0.18   |   0.8
Marginal | 0.7 |    0.3    |    1


Q6. A survey was conducted among 100 patients about smoking status. The following is the sample size split by smoking status (Yes or No) and gender (Male or Female). 


Table    | Smoking (Yes) | Smoking(No) | Total
---------|---------------|-------------|----------
Male     |       19      |     36      | 55
Female   |       13      |     32      | 45
Total    |       32      |     68      | 100 


The probability that a randomly selected patient is a male who smokes is 0.19.  

(a) Fill in all the elements of the table
(Ans) The table above is filled as specified.

(b) What is the probability of a randomly selected patient being a female?    
(Ans) P(female)= 45/100 = .45

(c) What is the probability of a randomly selected patient being a smoker?     
(Ans) P(smoker) = 32/100 = .32

(d)  What is the probability of a randomly selected smoker being a female?    
(Ans) P(female|smoker) = 13/32 = .40

#### Section 2: Data Analysis using R: Total points 25


Q1 : Using the dataset provided ("sample_patient_dataset.csv"), the task to build a 2x2 table for the studying the association between age at admission >70 and cardiac arrests. You can either use the sample table given below or build your own. Rememer to output both count and % in the table. Be sure to round the % to the nearest integer (e.g, 0.674 will be 67% and 0.675 will be 68%, see notes in Lecture2 on summary statistics as example). Fill in the code in the shaded areas.


Table         | Cardiac Arrests (Yes) | Cardiac Arrests (No) | Total
--------------|-----------------------|----------------------|----------
Age > 70 (%)  |       453   (2)       |     4728 (20)     | 5181 (21)
Age <= 70 (%) |       1672  (7)       |     17254(72)     | 18926 (79)
Total         |       2125   (9)     |     21982 (91)    | 24,107


```{r}

#library(readr)
library(dplyr)
library(lubridate)
library(formattable)

pdata <-  read.csv("/Users/shaileshdudala/R Projects/sample_patient_dataset.csv")
pd    <-  pdata %>% mutate(dob = mdy(dob), hosp_admission = mdy(hosp_admission)) %>% 
          mutate(age_at_admit = interval(dob, hosp_admission)/ dyears(1)) %>%
          select(c("age_at_admit", "had_cardiac_arrests")) 

card_70plus    <-  pd %>% filter(age_at_admit > 70 & had_cardiac_arrests == 1) %>% nrow()
nocard_70plus  <-  pd %>% filter(age_at_admit > 70 & had_cardiac_arrests == 0) %>% nrow()
card_under70   <-  pd %>% filter(age_at_admit < 70 & had_cardiac_arrests == 1) %>% nrow()
nocard_under70 <-  pd %>% filter(age_at_admit < 70 & had_cardiac_arrests == 0) %>% nrow()
totCard        <-  pd %>% filter(had_cardiac_arrests == 1) %>% nrow()
totNoCard      <-  pd %>% filter(had_cardiac_arrests == 0) %>% nrow()
tot70Plus      <-  pd %>% filter(age_at_admit > 70) %>% nrow()
totUnder70     <-  pd %>% filter(age_at_admit < 70) %>% nrow()
totPat         <- totCard + totNoCard
# Patients with Cardiac Arrests above 70
card_70plus    <-  pd %>% filter(age_at_admit > 70 & had_cardiac_arrests == 1) %>% nrow()
cat("Patients with Cardiac Arrests above 70: ",card_70plus)
x <- round(percent(card_70plus/totPat), digits = 2)
print(x)
#Patients wihout Cardiac Arrests above 70
nocard_70plus  <-  pd %>% filter(age_at_admit > 70 & had_cardiac_arrests == 0) %>% nrow()
cat("Patients wihout Cardiac Arrests above 70: ",nocard_70plus)
y <- round(percent(nocard_70plus/totPat), digits = 2)
print(y)
#Patients with Cardiac Arrests under 70
card_under70   <-  pd %>% filter(age_at_admit < 70 & had_cardiac_arrests == 1) %>% nrow()
cat("Patients wih Cardiac Arrests under 70: ",card_under70)
z <- round(percent(card_under70/totPat), digits = 2)
print(z)
#Patients without Cardiac Arrests under 70
nocard_under70 <-  pd %>% filter(age_at_admit < 70 & had_cardiac_arrests == 0) %>% nrow()
cat("Patients without Cardiac Arrests under 70: ",nocard_under70)
a <- round(percent(nocard_under70/totPat), digits = 2)
print(a)
#Total Patients with Cardiac Arrests
totCard        <-  pd %>% filter(had_cardiac_arrests == 1) %>% nrow()
cat("Total Patients with Cardiac Arrests: ",totCard)
b <- round(percent(totCard/totPat), digits = 2)
print(b)
#Patients without Cardiac Arrests
totNoCard      <-  pd %>% filter(had_cardiac_arrests == 0) %>% nrow()
cat("Total Patients without Cardiac Arrests: ", totNoCard)
c <- round(percent(totNoCard/totPat), digits = 2)
print(c)
#Total PAtients above age 70
tot70Plus      <-  pd %>% filter(age_at_admit > 70) %>% nrow()
cat("Total PAtients above age 70: ", tot70Plus)
d <- round(percent(tot70Plus/totPat), digits = 2)
print(d)
#Total Patients under age 70
totUnder70     <-  pd %>% filter(age_at_admit < 70) %>% nrow()
cat("Total Patients under age 70: ", totUnder70)
e <- round(percent(totUnder70/totPat), digits = 2)
print(e)
#Total Number of Patients
totPat <- tot70Plus + totUnder70
cat("The total number of Patients is ", totPat)





```



Q2: Create your own de-identified version of "patient_dataset.csv". Upload your de-identified dataset onto Canvas and write the de-identification code below. You will need to refer to the document "Deidentification.pdf" (on Canvas, look under files -> lectures -> lecture_2).


```{r}
#Import dataset
deid.in <- read.csv("/Users/shaileshdudala/R Projects/patient_dataset.csv")
# create random Ids
all.patients <- deid.in %>%
  select(patient.ids) %>%
  unique()
all.patients$random_id <- sample(nrow(all.patients), replace = FALSE)

# merge random ids to data frame
deid.in <- merge(deid.in, all.patients, by = "patient.ids")

# remove patient ids
deid.in <- deid.in %>%
select(-c(patient.ids))

#remove names
deid.in <- deid.in %>%
select(-c(patient.names))

# remove other identifiers
deid.in <- deid.in %>%
select(-c(street_address, city, zip_code, contact_number))

# First, format date of discharge
deid.in$hosp_admission_format  <- mdy(deid.in$hosp_admission, tz = "America/Chicago")

# Create new date-time variable for date of discharge
deid.in$hosp_discharge_format <- mdy(deid.in$hosp_discharge, tz = "America/Chicago")

# Using mutate from plyr and dplyr for data frame operations
deid.in <- mutate(deid.in,
              hosp_admission_format = mdy(deid.in$hosp_admission), 
              hosp_discharge_format = mdy(deid.in$hosp_discharge))

# Calculate diffreence between admission and discharge
deid.in <- mutate(deid.in,
               temp_interval2 = interval(hosp_admission_format, hosp_discharge_format))
# Calculate age
deid.in <- deid.in %>% mutate(dob_form =mdy(deid.in$dob, tz = "America/Chicago")) %>% mutate(temp_interval =interval(dob_form, hosp_admission_format)) %>% mutate(age_at_admit = temp_interval/ dyears(1))

# Interval to days
deid.in <-mutate(deid.in,length_of_stay= temp_interval2 /ddays(1))

# remove identifiers and interval
deid.in <- deid.in %>% select(-c(dob, hosp_admission, hosp_admission_format, hosp_discharge, hosp_discharge_format, temp_interval2, admitting_provider, dob_form, temp_interval))

# write document
write.csv(deid.in, "/Users/shaileshdudala/R Projects//patient_dataset_deidentified.csv",
row.names = FALSE, quote = TRUE)


```

