---
title: "NET"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}


run_sim = function(lam, p, d_avg) {
  # the population size
  N = 70

# the needle sharing connection
# probability as a function of the

q = d_avg/(N - 1)


# the transmission rate beta
bet = lam * p

# max time: one year, measured in days
tmax = 365

# set the infection indicator
y = rep(0, N)

# we will set person 1 to be infected at
# baseline
y[1] = 1

# set the infection time
ts = rep(NA, N)
ts[1] = 0

# set up a symmetric zero-diagonal
# contact matrix
A = matrix(0, nrow = N, ncol = N)
for (i in 1:(N - 1)) {
  for (j in (i + 1):N) {
    A[i, j] = rbinom(1, 1, q)
    A[j, i] = A[i, j]
  }
}
A[diag(A)] = 0

tcur = 0

while (tcur < tmax && any(y == 0)) {
  
  # the number of edges connected to each
  # susceptible unit that are attached to
  # an infected unit
  susc_edges = (1 - y) * (A %*% y)
  if (sum(susc_edges) == 0) 
    break
  
  # the waiting time to the next infection
  w = rexp(1, bet * sum(susc_edges))
  
  # if this would put us over the time
  # horizon, stop
  if (tcur + w > tmax) 
    break
  
  # choose the id of the next infected unit
  idx = sample(1:N, 1, prob = susc_edges)
  
  # mark this unit infected
  y[idx] = 1
  
  # record the infection time of the newly
  # infected unit
  ts[idx] = tcur + w
  
  # update the current time
  tcur = tcur + w
}

ts[y == 0] = Inf


return(sum(y))
#plot(sort(ts), 1:N, bty = "n", xlab = "Infection time")



plot_network_infections = function(t, lay) {
  set.seed(1)
  g = network(A, vertex.attr = list(y = ifelse(ts <= 
                                                 t, 1, 0)), matrix.type = "adjacency", 
              directed = FALSE)
  plot(g, vertex.col = ifelse(get.vertex.attribute(g, 
                                                   "y"), "red", "white"), vertex.cex = 2, 
       main = paste("t =", round(t, digits = 2)))
}


tseq = seq(0, max(ts[ts < Inf]), length.out = 9)
par(mfrow = c(3, 3), mar = c(1, 0, 1, 0))

#for (t in tseq) {
  #plot_network_infections(t, lay)
#}


}




```

```{r}

for(i in 1:100){
  run_sim(1, 0.01, 3)
  z[i] = sum(y==1)
  
}

hist(z, breaks = 10 )
```


1. Before doing any programming, look at the simulation code above. It is very concise, and runs efficiently on a computer, even when the network consists of hundreds of individual. Why? Go line by line through the while loop to understand each part. Which part chooses the next infected individual? Why are there no rate parameters involved in this choice?

2. Using the default values of input parameters, what is the average number of infections that occur in the population over one year?


3. Public health authorities attempt to help injection drug users stop using drugs by providing opioid agonist therapy (OAT) free of charge, thereby reducing the frequency λ of drug use and needle sharing. Assess the impact of reducing the frequency of drug use and needle sharing by plotting the distribution of total infections by λ as it ranges from 0 to 1.

```{r}
lambda = seq(0.02,2, by=0.02 )
X <- vector(mode="integer", length=length(lambda))

for(i in 1:length(lambda)){
  
  X[i] = mean(replicate(100, run_sim(lambda[i], 0.01, 3)))
  
}

qplot(x = lambda, y= X)
```


4. Public health authorities decide that they cannot limit injection drug use, but they can supply drug users with clean needles, thereby reducing the average number of needle sharing partners they have. Assess the impact of reducing the average degree by plotting the distribution of total infections by the average degree as it ranges from 1 to 10.

```{r}
deg = seq(1, 10, by=1 )
X <- vector(mode="integer", length=length(lambda))

for(i in 1:length(lambda)){
  
  X[i] = mean(replicate(100, run_sim(1, 0.01, deg[i])))
  
}


plot(x = lambda, y= X)
```


5. Drug use epidemiologists criticize the model described above because it assumes that all individuals are the same in their drug use characteristics, and differ only in their location in the network. You decide to remedy this by defining the needle sharing rate λi for susceptible individuals as a function of individualistic covariates. You decide to model
λi=exp[zi]
where
zi∼Normal(0,σ2)
You will need to rewrite the part of the simulation code that defined λ and the part that computes the waiting time to the next infection. Assess the impact of this individualistic variation by plotting total infections by σ2 as it ranges from 0 to 3.


```{r}

```

6. Extra Credit: Infectious disease epidemiologists criticize the model above because it assumes that individuals are infectious forever after contracting HCV. In reality, individuals may be diagnosed and treated with effective drugs that cure the infection. Rewrite this model as an agent-based stochastic SIR model, and assess the total number of infections as a function of the “removal rate” γ.

```{r}

```

